Does Anybody Really Know What Time It Is?

### 2015-07-04 intro.tex
More transferring from Word into LaTeX

>>>>

\subsubsection{Parsing}
One of the oldest data preparation problems observed in Tableau is the parsing of date scalars so that users can perform these kinds of analyses. Training materials from the early days of the company included examples of how to convert columns of integers in the form yyyyMMdd to date scalars. The documented solution at the time was to convert the integer to a string and perform some locale-dependent string operations before casting the string back to a date. Unfortunately, this approach had a number of problems:
\begin{itemize}
\item String operations are notoriously slow compared to scalar operations (typically 10-100x slower)
\item Default parsing of date formats is locale-dependent, and may not work when the workbook is shared across an international organization (e.g. between the US and European offices)
\item The expression code was hard to understand and maintain because it used a verbose, general-purpose string-handling syntax instead of a domain language.
\end{itemize}

This is but a single format. Our studies of the workbooks on Public suggest that there are hundreds of distinct temporal date formats in user data sets. Some are common, but others can be quite idiosyncratic. Table 1 shows a selection of unusual date formats found on Public.

% INSERT TABLE HERE

Our first solution to this problem was to add a new function to the Tableau calculation language called DATEPARSE. This function would take a string column and convert it to a datetime using a special purpose domain language for describing dates. Such functions exist in a number of databases (e.g. Oracle, Postgres and MySQL) and adding it to the TDE was straightforward, so it was a natural addition to the function library.
We chose the date parsing syntax defined by the International Components for Unicode (or ICU) project [8] for the common domain language because it mirrored what was available in the Tableau code base, and translating this syntax into the formats used by various database vendors (e.g. MySQL, Oracle, Postgres) was relatively painless. There are a few patches of non-overlapping functionality, but they tend to be obscure and we provide warnings when functionality is missing.

\subsubsection{	Usability}
After shipping this new function to our user base, we investigated how it was being applied in the field. We examined a few months of workbooks that had been uploaded to Public after the release to see if and how it was being used. We found that although there were a number of new calculations using DATEPARSE, the error rate for the domain language syntax was about 15\%. DATEPARSE was capable of solving the problem, and some users were able to discover it, but the syntax did not appear to be easy to use reliably.

\subsection{Automated Format Extraction}
One solution might have been to design a graphical environment that enabled users to construct valid patterns. This would have involved a substantial development investment with no guarantee that the result would be correct if the user misunderstood the environment. Instead, we have developed two algorithms for automating the derivation of the format string, each of which uses a different machine learning technique. Both algorithms result in over 90\% parsing accuracy – and 100\% syntactic correctness because they are machine generated.

% vidya: this probably should be its own section

\subsection{Related Work}
Data preparation has been a known analytic bottleneck since at least the description of the Potter’s Wheel system [4]. Since then, several other interactive data preparation systems have been proposed [5,6]. While effective, these systems all make assumptions about possible date formats (e.g. the domain system in [4]), which we suggest are too restrictive for real world data.
The Minimum Descriptive Length technique was first proposed in [7]. We have built on the version described in [4].
Related work on parsing languages, outlier detection.

\subsection{Overview}
The rest of this paper is organized as follows: The next section introduces the parameters of the problem space. The following two sections describe the two different algorithms, one using Minimum Descriptive Length and the other using Natural Language Processing. In section 5, we evaluate the algorithms on a corpus of 30K columns, both by sampling the outputs manually and then by using the algorithms to validate each other. We then discuss future work in section 6 and conclude in section 7.
